---
title: "hCBO compare protocols using cerebellar markers"
author: "Qiang (Danny) Chen"
date: "`r Sys.Date()`"
output:
  html_document:
    keep_md: true
    highlight: tango
    code_folding: hide
    number_sections: yes
    theme: united
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
  word_document:
    toc: yes
  pdf_document:
    toc: yes
  bibliography: biblio.bib
vignette: |
  %\VignetteIndexEntry{RNASeq} %\VignetteEngine{knitr::rmarkdown} %\VignetteEncoding{UTF-8} %\usepackage[utf8]{inputenc}
---

We integrated SCTL samples at day60 in our 2 month-old hCBO data with the following two datasets and identified DEGs among cerebellar markers across protocols. 

For the quadrato lab, we focused on three samples below (they generated three replicates at day 60):  
[GSE247974](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE247974)
GSM7904010: organoid D1, scRNA-seq  
GSM7904011: organoid D2, scRNA-seq  
GSM7904012: organoid D3, scRNA-seq  

For the pasca lab, we only focused on one sample (at day 72) below:  
[GSE233574](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE233574)
GSM7430706: OrganoidScreen; sample11; FGF2-50  

There is only one sample (GSM7430706) in GSE233574. We did sub-sampling in cells of GSM7430706 and calculated psudo-bulk of sub-samples for comparisons. 


```{r setup, echo=T, results="hide", message=F}
# knitr::opts_chunk$set(tidy=FALSE, cache=FALSE, echo=TRUE, dev="png", message=FALSE, error=FALSE, warning=FALSE)
knitr::opts_chunk$set(echo=TRUE, warning=F, message=F)

library("tidyr")
library("dplyr")
library("ggplot2")
library("knitr")
library("pheatmap")
library("DT")

library("Seurat")
library("DESeq2")

rm(list=ls())

d_meta = "/data/qiangchen/Projects/NCATS/SCTL/SeungMiRyu/hCBO/Data/"
d_proj = "/data/qiangchen/Projects/NCATS/SCTL/SeungMiRyu/hCBO/SCTLday60_GSE247974_GSE233574/"
d_dat  = paste0(d_proj, "Data/")
d_res  = paste0(d_proj, "Results/")


```

# Load merged data

```{r load, eval=T, echo=T, results='hide'}

f_count = paste0(d_dat, "count_matrix.SCTLday60_GSE247974_GSE233574.qced.rda")
load(f_count)

sctl_samples = sampleInfo[sampleInfo$Protocol=="SCTLday60", "ID"]

f_cb = paste0(d_meta, "Full_Brain_Region_Marker_Table_Ryu_et_al.csv")
gene_data = read.csv(f_cb)
selected_genes = unique(gene_data$MarkerGene)
print(length(selected_genes))

matched = selected_genes %in% rownames(count_matrix)
selected_genes = selected_genes[matched]
print(length(selected_genes))


```

# DEG

We compared protocols using 110 cerebellar markers. Among them, 106 genes were found in the merged data. 

```{r deg, eval=T, echo=T, collapse=T}

rownames(sampleInfo) = sampleInfo$ID

dds = DESeqDataSetFromMatrix(
    countData = count_matrix,
    colData = sampleInfo,
    design = ~ Protocol
)

dds = DESeq(dds, quiet=T)

f_dds = paste0(d_dat, "deg_model.SCTLday60_GSE247974_GSE233574.rds")
saveRDS(dds, f_dds)

norm_counts = counts(dds, normalized = TRUE)
rownames(norm_counts) = rownames(count_matrix)

norm_counts = norm_counts[selected_genes, ]
print(dim(norm_counts))

f_norm = paste0(d_dat, "norm_counts.SCTLday60_GSE247974_GSE233574.CerebellarGenes.rds")
saveRDS(norm_counts, f_norm)


```

# DEG: SCTLday60 vs GSE247974

```{r contrast1, echo=T, message=T, warning=F, collapse=T}

deg = results(dds, contrast=c("Protocol", "SCTLday60", "GSE247974"))

deg = as.data.frame(deg)
deg = cbind(Gene = rownames(deg), deg)

colnames(deg) = c("Gene", "baseMean", "log2FC", "lfcSE", "stat", "pval", "padj")
deg = deg[order(deg$padj),]

f_deg = paste0(d_res, "deg.SCTLday60_vs_GSE247974.tsv")
write.table(deg, f_deg, sep="\t", row.names=F, quote=F)

deg.cb = deg[selected_genes, ]
deg.cb = deg.cb[order(deg.cb$padj), ]

f_deg = paste0(d_res, "deg.SCTLday60_vs_GSE247974.cerebellar_genes.tsv")
write.table(deg.cb, f_deg, sep="\t", row.names=F, quote=F)

datatable(
    deg.cb,
    caption = "SCTLday60 vs GSE247974",
    rownames = FALSE,
    options = list(
        dom = 'Bfrtip', # B for buttons, frtip for other table features
        buttons = c('csv', 'excel'), # Add CSV and Excel download buttons
        pageLength = 10
    )    
) %>% 
    formatRound(columns = c("baseMean", "log2FC", "lfcSE", "stat", "pval", "padj"), digits = 2)

deg.cb.pos = deg.cb[deg.cb$log2FC>0, ]
deg.cb.pos = deg.cb.pos[order(deg.cb.pos$log2FC, decreasing=T),]

deg.cb.pos.top20 = deg.cb.pos[1:20, ]

expr.top20 = norm_counts[deg.cb.pos.top20$Gene, c(which(colnames(norm_counts) %in% sctl_samples),
    grep("^GSE247974", colnames(norm_counts)))]

pheatmap(
    expr.top20,
    scale = "row",           
    main = "Top 20 genes with expression in SCTLday60 higher than in GSE233574 (sorted by log2FC)",
    fontsize_row = 10,   
    fontsize_col = 10,   
    show_rownames = TRUE,
    show_colnames = TRUE,
    cluster_cols = FALSE
)

deg.cb.neg = deg.cb[deg.cb$log2FC<=0, ]
deg.cb.neg = deg.cb.neg[order(deg.cb.neg$log2FC, decreasing=F),]

deg.cb.neg.top20 = deg.cb.neg[1:20, ]

expr.top20 = norm_counts[deg.cb.neg.top20$Gene, c(which(colnames(norm_counts) %in% sctl_samples),
    grep("^GSE247974", colnames(norm_counts)))]

pheatmap(
    expr.top20,
    scale = "row",           
    main = "Top 20 genes with expression in SCTLday60 lower than in GSE233574 (sorted by log2FC)",
    fontsize_row = 10,   
    fontsize_col = 10,   
    show_rownames = TRUE,
    show_colnames = TRUE,
    cluster_cols = FALSE
)


```

# DEG: SCTLday60 vs GSE233574

```{r contrast2, echo=T, message=T, warning=F, collapse=T}

deg = results(dds, contrast=c("Protocol", "SCTLday60", "GSE233574"))

deg = as.data.frame(deg)
deg = cbind(Gene = rownames(deg), deg)

colnames(deg) = c("Gene", "baseMean", "log2FC", "lfcSE", "stat", "pval", "padj")
deg = deg[order(deg$padj), ]

f_deg = paste0(d_res, "deg.SCTLday60_vs_GSE233574.tsv")
write.table(deg, f_deg, sep="\t", row.names=F, quote=F)

deg.cb = deg[selected_genes, ]
deg.cb = deg.cb[order(deg.cb$padj), ]

f_deg = paste0(d_res, "deg.SCTLday60_vs_GSE233574.cerebellar_genes.tsv")
write.table(deg.cb, f_deg, sep="\t", row.names=F, quote=F)

datatable(
    deg.cb,
    extensions = 'Buttons',
    caption = "SCTLday60 vs GSE233574",
    rownames = FALSE,
    options = list(
        dom = 'Bfrtip', # B for buttons, frtip for other table features
        buttons = c('csv', 'excel'), # Add CSV and Excel download buttons
        pageLength = 10
    )    
) %>% 
    formatRound(columns = c("baseMean", "log2FC", "lfcSE", "stat", "pval", "padj"), digits = 2)

deg.cb.pos = deg.cb[deg.cb$log2FC>0, ]
deg.cb.pos = deg.cb.pos[order(deg.cb.pos$log2FC, decreasing=T),]

deg.cb.pos.top20 = deg.cb.pos[1:20, ]

expr.top20 = norm_counts[deg.cb.pos.top20$Gene, c(which(colnames(norm_counts) %in% sctl_samples),
    grep("^GSE233574", colnames(norm_counts)))]

pheatmap(
    expr.top20,
    scale = "row",           
    main = "Top 20 genes with expression in SCTLday60 higher than in GSE233574 (sorted by log2FC)",
    fontsize_row = 10,   
    fontsize_col = 10,   
    show_rownames = TRUE,
    show_colnames = TRUE,
    cluster_cols = FALSE
)

deg.cb.neg = deg.cb[deg.cb$log2FC<=0, ]
deg.cb.neg = deg.cb.neg[order(deg.cb.neg$log2FC, decreasing=F),]

deg.cb.neg.top20 = deg.cb.neg[1:20, ]

expr.top20 = norm_counts[deg.cb.neg.top20$Gene, c(which(colnames(norm_counts) %in% sctl_samples),
    grep("^GSE233574", colnames(norm_counts)))]

pheatmap(
    expr.top20,
    scale = "row",           
    main = "Top 20 genes with expression in SCTLday60 lower than GSE233574 (sorted by log2FC)",
    fontsize_row = 10,   
    fontsize_col = 10,   
    show_rownames = TRUE,
    show_colnames = TRUE,
    cluster_cols = FALSE
)


```

# Session Information

```{r info, echo=T, message=T, warning=F, collapse=T}

sessionInfo()

rm(list=ls())


```

